\pagestyle{empty}
%% copyrightpage
\begingroup
\footnotesize
\parindent 0pt
\parskip \baselineskip
\textcopyright{} 2016 Antonius Torode \\
All rights reserved.

This work may be distributed and/or modified under the conditions of Antonius' General Purpose License.

The Original Maintainer of this work is: Antonius Torode.

The Current Maintainer of this work is: Antonius Torode.

This document is designed for the sole purpose of providing documentation for the Generations of Nuclear Atrophy (GINA) program. GINA is a program written for and created at the National Superconducting Cyclotron Laboratory (NSCL). Throughout this document, ``GINA" will be used as a reference to the ``Generations of Nuclear Atrophy" program. This acronym is solely designed for use in conjunction with the program and was originally created by the creator of this document. GINA is designed for specific purposes at the NSCL. It can be adapted to work in other situations and for alternate purposes. GINA is not be to used for gaining profit. GINA may be freely used and adapted to be used solely at Michigan State University and the NSCL.

This document is continually in development as GINA is. The version number of this document is to correspond with the version of GINA it covers.


\begin{center}
\begin{tabular}{ll}
Most Current Revision Date: &  \today 
\end{tabular}
\end{center}

\vfill

Torode, A.\\
\hspace*{2em} National Superconducting Cyclotron Laboratory. \\
\hspace*{2em} 2017, Undergraduate Researcher. \\
\hspace*{2em} Includes Source Code and References \\
\hspace*{2em} ISBN: NONE \\
\hspace*{2em} Book designed in \LaTeX



\endgroup
\clearpage